---
title: "distance_moved"
---

```{r load libraries}
# Load required libraries
library(dplyr)
library(geosphere)
```

```{r read in the data}
# Read in the data
fish_flow_receiver_stream_corrected <- read.csv("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/radio_tracking_2024/data/combined_data/fish_flow_receiver_stream_corrected.csv")

head(fish_flow_receiver_stream_corrected)
```

```{r distance moved 1.0}

# Convert `datetime` to POSIXct
fish_flow_receiver_stream_corrected <- fish_flow_receiver_stream_corrected %>%
  mutate(trackedTime_EST = as.POSIXct(trackedTime_EST, format = "%Y-%m-%d %H:%M:%S"))

# Select iPad points or receiver with highest power when no iPad exists for the day/shift
filtered_data <- fish_flow_receiver_stream_corrected %>%
  group_by(radioID, date, shift) %>%
  filter(
    (source == "iPad") | # Keep iPad points
    (!any(source == "iPad") & source == "receiver" & power == max(power, na.rm = TRUE)) # Receiver only if no iPad
  ) %>%
  ungroup()

# Calculate distances
filtered_data <- filtered_data %>%
  arrange(radioID, trackedTime_EST) %>% # Ensure proper order
  group_by(radioID) %>%
  mutate(
    # Add debugging print statements
    distance_m = {
      if (n() > 1) {
        print(paste("Calculating distance for radioID:", first(radioID))) # Debugging info
        c(
          NA, # First point has no previous point
          distHaversine(
            cbind(lon[-n()], lat[-n()]), # Previous points
            cbind(lon[-1], lat[-1])      # Next points
          )
        )
      } else {
        print(paste("Skipping radioID:", first(radioID), "as it has only one point")) # Debugging
        NA_real_
      }
    }
  ) %>%
  ungroup()

# View results
print(filtered_data)

```




```{r distance moved 2.0}

# Convert `datetime` to POSIXct
fish_flow_receiver_stream_corrected <- fish_flow_receiver_stream_corrected %>%
  mutate(trackedTime_EST = as.POSIXct(trackedTime_EST, format = "%Y-%m-%d %H:%M:%S"))

# Select iPad points or receiver with highest power when no iPad exists for the day/shift
#filtered_data <- fish_flow_receiver_stream_corrected %>%
  #group_by(radioID, date, shift) %>%
  #filter(
    #(source == "iPad") | # Keep iPad points
    #(!any(source == "iPad") & source == "receiver" & power == max(power, na.rm = TRUE)) # Receiver only if no #iPad
  #) %>%
  #ungroup()

# Identify all day+shift combinations that contain "OUT OF WATER"
out_of_water_days <- fish_flow_receiver_stream_corrected %>%
  filter(status == "OUT OF WATER") %>%
  select(radioID, date, shift) %>%
  distinct()

# Filter out points from any day+shift where "OUT OF WATER" occurs, while keeping "RECOVERED" points
filtered_data <- filtered_data %>%
  anti_join(out_of_water_days, by = c("radioID", "date", "shift")) %>%
  group_by(radioID, date, shift) %>%
  filter(
    (source == "iPad") | # Keep iPad points
    (!any(source == "iPad") & source == "receiver" & power == max(power, na.rm = TRUE)) # Receiver only if no iPad
  ) %>%
  ungroup()

# Calculate distances
filtered_data <- filtered_data %>%
  arrange(radioID, trackedTime_EST) %>% # Ensure proper order
  group_by(radioID) %>%
  mutate(
    # Add debugging print statements
    distance_m = {
      if (n() > 1) {
        print(paste("Calculating distance for radioID:", first(radioID))) # Debugging info
        c(
          NA, # First point has no previous point
          distHaversine(
            cbind(lon[-n()], lat[-n()]), # Previous points
            cbind(lon[-1], lat[-1])      # Next points
          )
        )
      } else {
        print(paste("Skipping radioID:", first(radioID), "as it has only one point")) # Debugging
        NA_real_
      }
    }
  ) %>%
  ungroup()

# View results
print(filtered_data)

```



##################################################

# Karli RiverDist Script

## Load your libraries (You might not need them all right now)

```{r setup libraries}
# Required packages for script
require(tidyverse) # Used for Data Wrangling and Clean-up
require(data.table) # Used for Data Wrangling and Clean-up
require(riverdist) # Used for Individual Movement
require(RMark) # Used for Mark-Recapture Data (Survival and Detection Rates)
library(readxl)
library(gridExtra)
library(moments)

# Convert points to UTMs from decimal degrees
library(sf)
#library(rgdal) # package no longer exists

# For the coordinates function???
library(sp)
```
## Load dataset

```{r load and filter fish data}
fish_flow_receiver_stream_corrected <- read.csv("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/radio_tracking_2024/data/combined_data/fish_flow_receiver_stream_corrected.csv")

# Select iPad points or receiver with highest power when no iPad exists for the day/shift
fish_data_filtered <- fish_flow_receiver_stream_corrected %>%
  group_by(radioID, date, shift) %>%
  mutate(has_iPad = any(source == "iPad"), 
         has_receiver = any(source == "receiver")) %>%  # Check if any receiver points exist
  filter(
    (has_iPad & source == "iPad") |  # Keep iPad points if they exist
    (has_receiver & !has_iPad & source == "receiver" & power == max(power, na.rm = TRUE))  # Only filter receiver points with max power if receiver exists
  ) %>%
  ungroup()
```


## Underhill Brook

1. Load your flowline shapefiles (do this for each watershed separately)

```{r Load in watershed flowlines}

underhill_flowlines <- lines <- st_read("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_UnderhillBrook.shp") # Reads shapefile, these shapefiles are already in geographic coordinate system NAD83
st_crs(underhill_flowlines) # This just checks to make sure that your lines are in GCS NAD83
underhill_flowlines <- st_transform(underhill_flowlines, 26918) # Projects to UTM Zone 18N (the code for this is 26918)

## Please read note before running next line! Comment the next line out if you already reprojected the raw shapefile.
#st_write(underhill_flowlines, "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_UnderhillBrook_UTMZONE18N.shp") #Once you run this, it will add a new shapefile to your "shapefiles" folder! It will give you a warning if it's already done. I did it for underhill already, but you'll need to repeat this process for each watershed. 


underhill_flowlines_UTMZONE18N <- line2network(path = "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles", # Points to the folder in your R Project folder
                                    layer = "NHDFlowline_UnderhillBrook_UTMZONE18N", # Identifies the shapefile you want to pull from folder
                                    tolerance = 10) # This is the spatial tolerance used to determine connectivity between line segments) 


plot(x=underhill_flowlines_UTMZONE18N) # checks imported line
topologydots(rivers = underhill_flowlines_UTMZONE18N) # checks line topology

# Setting the MOUTH of your watershed allows you to look at directional movement. The "seg = " value will be the segment that includes your mouth, which you can observe in the plot function above. In Underhill, it's segment 46. The "vert = " value identifies the correct node of that segment. Using "showends" on the segment you want will show you what vertex to select.
plot(x=underhill_flowlines_UTMZONE18N)
showends(seg = 46, rivers = underhill_flowlines_UTMZONE18N) # identifies the true mouth of your watershed
underhill_mouth <- setmouth(seg = 46, vert = 4, rivers = underhill_flowlines_UTMZONE18N) # sets the mouth to the segment and vertex you need.

```

2. Load your fish points for the watershed you are analyzing (do this for each watershed separately)
```{r Load in fish points}

# read in your fish points and filter by the watershed you're looking at. Make sure to have X and Y columns available in decimal degrees

#fish_flow_receiver_stream_corrected <- read.csv("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/radio_tracking_2024/data/combined_data/fish_flow_receiver_stream_corrected.csv")

# Keep only points from "UNDERHILL" and also remove rows where lat or lon are NA
underhill_points <- fish_data_filtered %>%
  filter(river == "UNDERHILL" & !is.na(lon) & !is.na(lat))

# Make Lat Lon columns numeric

underhill_points_spatial <- underhill_points %>% 
  mutate(POINT_Y = as.numeric(lat)) %>% # change to whatever your Y coord column is
  mutate(POINT_X = as.numeric(lon)) # change to whatever your X coord column is
coordinates(underhill_points_spatial) <- c("POINT_X","POINT_Y") # identifies the coordinate columns
proj4string(underhill_points_spatial) <- CRS("+proj=longlat +datum=WGS84") 
res <- spTransform(underhill_points_spatial, CRS("+proj=utm +zone=18 +ellps=WGS84")) # reprojects the points to UTM ZONE 18
as(res, "SpatialPoints")
res <- as.data.frame(res)
res <- res %>% 
  rename(POINT_X = coords.x1, POINT_Y = coords.x2) %>%  # Rename coords.x1 and coords.x2
  mutate(Rownumber =row_number())
underhill_segvert <- xy2segvert(x = res$POINT_X, y = res$POINT_Y, rivers = underhill_flowlines_UTMZONE18N) # change the "res$" to your X and Y coord columns
  
## Display raw points (red dots) and snapped points (blue). Run these three lines all at once
plot(x= underhill_flowlines_UTMZONE18N)
points(res$POINT_X, res$POINT_Y, pch=16, col="red") # shows raw points
riverpoints(seg = underhill_segvert$seg, vert = underhill_segvert$vert, rivers = underhill_flowlines_UTMZONE18N, pch = 15, col="blue")


```

```{r snapping distance}
hist(underhill_segvert$snapdist, main="snapping distance (m)")
```
```{r displaying point data in river locations}
zoomtoseg(seg=c(6, 44, 18), rivers= underhill_flowlines_UTMZONE18N)
points(res$POINT_X, res$POINT_Y, pch=16, col="red")
riverpoints(seg=underhill_segvert$seg, vert=underhill_segvert$vert, rivers= underhill_flowlines_UTMZONE18N, 
            pch=15, 
            col="blue")
```

```{r}

nrow(res)  # Total points before snapping
nrow(underhill_segvert)  # Total points successfully snapped





# Merge the snapped data with the original dataset using row numbers
underhill_segvert <- underhill_segvert %>%
  mutate(Rownumber = row_number()) # Assign row numbers if not already there

# Ensure `res` also has row numbers
#res <- res %>%
  #mutate(Rownumber = row_number())

# Merge snapped points back with the original dataset to retain radioID and trackedTime_EST
underhill_snapped <- left_join(res %>% select(Rownumber, radioID, trackedTime_EST), 
                               underhill_segvert, 
                               by = "Rownumber")

# Arrange data by fish ID and datetime to ensure chronological order
underhill_snapped <- underhill_snapped %>%
  arrange(radioID, trackedTime_EST)

duplicates <- underhill_snapped %>%
  group_by(radioID, trackedTime_EST) %>%
  filter(n() > 1)

print(duplicates)

# Compute distances using `riverdistance()`
underhill_snapped <- underhill_snapped %>%
  group_by(radioID) %>%
  mutate(
    network_distance = c(NA, # First point has no previous distance
      riverdistance(
        startseg = seg[-n()], startvert = vert[-n()], 
        endseg = seg[-1], endvert = vert[-1], 
        rivers = underhill_flowlines_UTMZONE18N
      )
    )
  ) %>%
  ungroup()

```


```{r computing network distances between sequential observations of individuals}
riverdistanceseq(unique=res$radioID, survey=res$trackedTime_EST, seg=res$seg, 
                   vert=res$vert, rivers=underhill_flowlines_UTMZONE18N)
```



### Dickey Brook

1. Load your flowline shapefiles (do this for each watershed separately)

```{r Load in watershed flowlines}

dickey_flowlines <- lines <- st_read("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_DickeyBrook.shp") # Reads shapefile, these shapefiles are already in geographic coordinate system NAD83
st_crs(dickey_flowlines) # This just checks to make sure that your lines are in GCS NAD83
dickey_flowlines <- st_transform(dickey_flowlines, 26918) # Projects to UTM Zone 18N (the code for this is 26918)

## Please read note before running next line! Comment the next line out if you already reprojected the raw shapefile.
#st_write(dickey_flowlines, "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_DickeyBrook_UTMZONE18N.shp") #Once you run this, it will add a new shapefile to your "shapefiles" folder! It will give you a warning if it's already done. I did it for underhill already, but you'll need to repeat this process for each watershed. 


dickey_flowlines_UTMZONE18N <- line2network(path = "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles", # Points to the folder in your R Project folder
                                    layer = "NHDFlowline_DickeyBrook_UTMZONE18N", # Identifies the shapefile you want to pull from folder
                                    tolerance = 10) # This is the spatial tolerance used to determine connectivity between line segments) 


plot(x=dickey_flowlines_UTMZONE18N) # checks imported line
topologydots(rivers = dickey_flowlines_UTMZONE18N) # checks line topology

# Setting the MOUTH of your watershed allows you to look at directional movement. The "seg = " value will be the segment that includes your mouth, which you can observe in the plot function above. In Underhill, it's segment 46. The "vert = " value identifies the correct node of that segment. Using "showends" on the segment you want will show you what vertex to select.
plot(x=dickey_flowlines_UTMZONE18N)
showends(seg = 42, rivers = dickey_flowlines_UTMZONE18N) # identifies the true mouth of your watershed
dickey_mouth <- setmouth(seg = 42, vert = 19, rivers = dickey_flowlines_UTMZONE18N) # sets the mouth to the segment and vertex you need.

```

2. Load your fish points for the watershed you are analyzing (do this for each watershed separately)
```{r Load in fish points}

# read in your fish points and filter by the watershed you're looking at. Make sure to have X and Y columns available in decimal degrees

fish_flow_receiver_stream_corrected <- read.csv("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/radio_tracking_2024/data/combined_data/fish_flow_receiver_stream_corrected.csv")

# Keep only points from "DICKEY" and also remove rows where lat or lon are NA
dickey_points <- fish_flow_receiver_stream_corrected %>%
  filter(river == "DICKEY" & !is.na(lon) & !is.na(lat))

# Make Lat Lon columns numeric

dickey_points_spatial <- dickey_points %>% 
  mutate(POINT_Y = as.numeric(lat)) %>% # change to whatever your Y coord column is
  mutate(POINT_X = as.numeric(lon)) # change to whatever your X coord column is
coordinates(dickey_points_spatial) <- c("POINT_X","POINT_Y") # identifies the coordinate columns
proj4string(dickey_points_spatial) <- CRS("+proj=longlat +datum=WGS84") 
res <- spTransform(dickey_points_spatial, CRS("+proj=utm +zone=18 +ellps=WGS84")) # reprojects the points to UTM ZONE 18
as(res, "SpatialPoints")
res <- as.data.frame(res)
res <- res %>% 
  rename(POINT_X = coords.x1, POINT_Y = coords.x2) %>%  # Rename coords.x1 and coords.x2
  mutate(Rownumber =row_number())
dickey_segvert <- xy2segvert(x = res$POINT_X, y = res$POINT_Y, rivers = dickey_flowlines_UTMZONE18N) # change the "res$" to your X and Y coord columns
  
## Display raw points (red dots) and snapped points (blue). Run these three lines all at once
plot(x= dickey_flowlines_UTMZONE18N)
points(res$POINT_X, res$POINT_Y, pch=16, col="red") # shows raw points
riverpoints(seg = dickey_segvert$seg, vert = dickey_segvert$vert, rivers = dickey_flowlines_UTMZONE18N, pch = 15, col="blue")


```
```{r snapping distance}
hist(dickey_segvert$snapdist, main="snapping distance (m)")
```

```{r displaying point data in river locations}
zoomtoseg(seg=c(42, 8, 83), rivers= dickey_flowlines_UTMZONE18N)
points(res$POINT_X, res$POINT_Y, pch=16, col="red")
riverpoints(seg=dickey_segvert$seg, vert=dickey_segvert$vert, rivers= dickey_flowlines_UTMZONE18N, 
            pch=15, 
            col="blue")
```

### Dry Brook

1. Load your flowline shapefiles (do this for each watershed separately)

```{r Load in watershed flowlines}

dry_flowlines <- lines <- st_read("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_DryBrook.shp") # Reads shapefile, these shapefiles are already in geographic coordinate system NAD83
st_crs(dry_flowlines) # This just checks to make sure that your lines are in GCS NAD83
dry_flowlines <- st_transform(dry_flowlines, 26918) # Projects to UTM Zone 18N (the code for this is 26918)

## Please read note before running next line! Comment the next line out if you already reprojected the raw shapefile.
#st_write(dry_flowlines, "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_DryBrook_UTMZONE18N.shp") #Once you run this, it will add a new shapefile to your "shapefiles" folder! It will give you a warning if it's already done. I did it for underhill already, but you'll need to repeat this process for each watershed. 


dry_flowlines_UTMZONE18N <- line2network(path = "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles", # Points to the folder in your R Project folder
                                    layer = "NHDFlowline_DryBrook_UTMZONE18N", # Identifies the shapefile you want to pull from folder
                                    tolerance = 10) # This is the spatial tolerance used to determine connectivity between line segments) 


plot(x=dry_flowlines_UTMZONE18N) # checks imported line
topologydots(rivers = dry_flowlines_UTMZONE18N) # checks line topology

# Setting the MOUTH of your watershed allows you to look at directional movement. The "seg = " value will be the segment that includes your mouth, which you can observe in the plot function above. In Underhill, it's segment 46. The "vert = " value identifies the correct node of that segment. Using "showends" on the segment you want will show you what vertex to select.
plot(x=dry_flowlines_UTMZONE18N)
showends(seg = 9, rivers = dry_flowlines_UTMZONE18N) # identifies the true mouth of your watershed
dry_mouth <- setmouth(seg = 9, vert = 19, rivers = dry_flowlines_UTMZONE18N) # sets the mouth to the segment and vertex you need.

```

2. Load your fish points for the watershed you are analyzing (do this for each watershed separately)
```{r Load in fish points}

# read in your fish points and filter by the watershed you're looking at. Make sure to have X and Y columns available in decimal degrees

fish_flow_receiver_stream_corrected <- read.csv("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/radio_tracking_2024/data/combined_data/fish_flow_receiver_stream_corrected.csv")

# Keep only points from "DICKEY" and also remove rows where lat or lon are NA
dry_points <- fish_flow_receiver_stream_corrected %>%
  filter(river == "DRY UPPER" & !is.na(lon) & !is.na(lat))

# Make Lat Lon columns numeric

dry_points_spatial <- dry_points %>% 
  mutate(POINT_Y = as.numeric(lat)) %>% # change to whatever your Y coord column is
  mutate(POINT_X = as.numeric(lon)) # change to whatever your X coord column is
coordinates(dry_points_spatial) <- c("POINT_X","POINT_Y") # identifies the coordinate columns
proj4string(dry_points_spatial) <- CRS("+proj=longlat +datum=WGS84") 
res <- spTransform(dry_points_spatial, CRS("+proj=utm +zone=18 +ellps=WGS84")) # reprojects the points to UTM ZONE 18
as(res, "SpatialPoints")
res <- as.data.frame(res)
res <- res %>% 
  rename(POINT_X = coords.x1, POINT_Y = coords.x2) %>%  # Rename coords.x1 and coords.x2
  mutate(Rownumber =row_number())
dry_segvert <- xy2segvert(x = res$POINT_X, y = res$POINT_Y, rivers = dry_flowlines_UTMZONE18N) # change the "res$" to your X and Y coord columns
  
## Display raw points (red dots) and snapped points (blue). Run these three lines all at once
plot(x= dry_flowlines_UTMZONE18N)
points(res$POINT_X, res$POINT_Y, pch=16, col="red") # shows raw points
riverpoints(seg = dry_segvert$seg, vert = dry_segvert$vert, rivers = dry_flowlines_UTMZONE18N, pch = 15, col="blue")


```

```{r snapping distance}
hist(dry_segvert$snapdist, main="snapping distance (m)")
```

### Amethyst Brook

1. Load your flowline shapefiles (do this for each watershed separately)

```{r Load in watershed flowlines}

amethyst_flowlines <- lines <- st_read("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_AmethystBrook.shp") # Reads shapefile, these shapefiles are already in geographic coordinate system NAD83
st_crs(amethyst_flowlines) # This just checks to make sure that your lines are in GCS NAD83
amethyst_flowlines <- st_transform(amethyst_flowlines, 26918) # Projects to UTM Zone 18N (the code for this is 26918)

## Please read note before running next line! Comment the next line out if you already reprojected the raw shapefile.
#st_write(amethyst_flowlines, "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles/NHDFlowline_AmethystBrook_UTMZONE18N.shp") #Once you run this, it will add a new shapefile to your "shapefiles" folder! It will give you a warning if it's already done. I did it for underhill already, but you'll need to repeat this process for each watershed. 


amethyst_flowlines_UTMZONE18N <- line2network(path = "C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/streamShapeFiles/Riverdist MA BKT/Shapefiles", # Points to the folder in your R Project folder
                                    layer = "NHDFlowline_AmethystBrook_UTMZONE18N", # Identifies the shapefile you want to pull from folder
                                    tolerance = 10) # This is the spatial tolerance used to determine connectivity between line segments) 


plot(x=amethyst_flowlines_UTMZONE18N) # checks imported line
topologydots(rivers = amethyst_flowlines_UTMZONE18N) # checks line topology

# Setting the MOUTH of your watershed allows you to look at directional movement. The "seg = " value will be the segment that includes your mouth, which you can observe in the plot function above. In Underhill, it's segment 46. The "vert = " value identifies the correct node of that segment. Using "showends" on the segment you want will show you what vertex to select.
plot(x=amethyst_flowlines_UTMZONE18N)
showends(seg = 404, rivers = amethyst_flowlines_UTMZONE18N) # identifies the true mouth of your watershed
amethyst_mouth <- setmouth(seg = 404, vert = 81, rivers = amethyst_flowlines_UTMZONE18N) # sets the mouth to the segment and vertex you need.

```

2. Load your fish points for the watershed you are analyzing (do this for each watershed separately)
```{r Load in fish points}

# read in your fish points and filter by the watershed you're looking at. Make sure to have X and Y columns available in decimal degrees

fish_flow_receiver_stream_corrected <- read.csv("C:/Users/jpilchik/OneDrive - DOI/Documents/My study plan/Tagging/Tracking/radio_tracking_2024/data/combined_data/fish_flow_receiver_stream_corrected.csv")

# Keep only points from "DICKEY" and also remove rows where lat or lon are NA
amethyst_points <- fish_flow_receiver_stream_corrected %>%
  filter((river == "AMETHYST" | river == "BUFFAM" | river == "HARRIS") & !is.na(lon) & !is.na(lat))

# Make Lat Lon columns numeric

amethyst_points_spatial <- amethyst_points %>% 
  mutate(POINT_Y = as.numeric(lat)) %>% # change to whatever your Y coord column is
  mutate(POINT_X = as.numeric(lon)) # change to whatever your X coord column is
coordinates(amethyst_points_spatial) <- c("POINT_X","POINT_Y") # identifies the coordinate columns
proj4string(amethyst_points_spatial) <- CRS("+proj=longlat +datum=WGS84") 
res <- spTransform(amethyst_points_spatial, CRS("+proj=utm +zone=18 +ellps=WGS84")) # reprojects the points to UTM ZONE 18
as(res, "SpatialPoints")
res <- as.data.frame(res)
res <- res %>% 
  rename(POINT_X = coords.x1, POINT_Y = coords.x2) %>%  # Rename coords.x1 and coords.x2
  mutate(Rownumber =row_number())
amethyst_segvert <- xy2segvert(x = res$POINT_X, y = res$POINT_Y, rivers = amethyst_flowlines_UTMZONE18N) # change the "res$" to your X and Y coord columns
  
## Display raw points (red dots) and snapped points (blue). Run these three lines all at once
plot(x= amethyst_flowlines_UTMZONE18N)
points(res$POINT_X, res$POINT_Y, pch=16, col="red") # shows raw points
riverpoints(seg = amethyst_segvert$seg, vert = amethyst_segvert$vert, rivers = amethyst_flowlines_UTMZONE18N, pch = 15, col="blue")


```



```{r snapping distance}
hist(amethyst_segvert$snapdist, main="snapping distance (m)")
```






